\bf Forecast reconciliation: A geometric view with new insights on bias correction

### 2020-01-29 Hierarchical-forecast-reconciliation/HierarchicalForecastReconciliation.tex
finished MinT proof
changes to discussion of loss functions and reconciliation

In this section we discuss two sensible properties that can be achieved by reconciliation via orthogonal projection.
\begin{itemize}
\item The first is that reconciliation should adjust the base forecasts as little as possible; i.e. the base and reconciled forecasts should be `close'.  {\color{blue} This property is particularly emphasised by \cite{NysEtAl2019}}
\item The second is that reconciliation in some sense should improve forecast accuracy, or more loosely, that the reconciled forecast should be `closer' to the realised value targeted by the forecast.  {\color{blue} This property is particularly emphasised by \cite{WicEtAl2019}.}
\end{itemize}

To address the first of these properties we make the concept of closeness more concrete by considering the Euclidean distance between the base forecast $\hat{\bm{y}}_{t+h|t}$ and the reconciled forecast $\tilde{\bm{y}}_{t+h|t}$. A property of an orthogonal projection is that the distance between $\hat{\bm{y}}_{t+h|t}$ and $\tilde{\bm{y}}_{t+h|t}$ is minimal over any possible $\tilde{\bm{y}}\in\mathfrak{s}$. {\color{blue} For Euclidean distance this can be expressed as minimising a loss function $\|\hat{\bm{y}}_{t+h|t}-\bm{S}\bm{G}\hat{\bm{y}}_{t+h|t}\|$ with respect to $\bm{G}$ where $\|.\|$ denotes the $L_2$ norm.}  In this sense reconciliation via orthogonal projection {\color{blue} $\bm{G}=(\bm{S}'\bm{S})^{-1}\bm{S}'$} leads to the smallest possible adjustments of the base forecasts.

The {\color{blue} second property is concerned with a loss function $L({\bm{y}}_{t+h},\breve{\bm{y}}_{t+h})=\|{\bm{y}}_{t+h}-\bm{S}\bm{G}\breve{\bm{y}}_{t+h|t}\|$ where $\breve{\bm{y}}_{t+h|t}$ is either a base or reconciled forecast.  More specifically, the aim is to reconcile so that $L({\bm{y}}_{t+h},\hat{\bm{y}}_{t+h})-L({\bm{y}}_{t+h},\tilde{\bm{y}}_{t+h})$ is guaranteed to be positive.\, that is} reconciled forecasts should always be closer to the target than base forecasts. {\color{blue} This idea was} touched upon in Section 2.3 of \citet{WicEtAl2019} for the {\color{blue} special} case of MinT\@ {\color{blue} which has a slightly different loss function}. Here we provide a new explicit {\color{blue} and general} proof of that result. We do so first in the case {\color{blue} of a loss function based on Euclidean distance} where the geometric intuition of the proof is clear and then generalise the result to reconciliation using any projection matrix in Section~\ref{sec:oblique}

@@

In other ways, Theorem~\ref{th:distred} is weaker than perhaps often understood. First, when improvements in forecast accuracy are discussed in the context of the theorem, this refers to a very specific measure of forecast accuracy. In particular, this measure is the square root of the sum of squared forecast errors of \emph{all} variables in the hierarchy. Consequently, while forecast improvement is guaranteed for the hierarchy overall, {\color{blue} forecast errors are treated equally across all series} and reconciliation can {\color{blue} still} lead to less accurate forecasts for individual series. Second, although orthogonal projections are guaranteed to improve on base forecasts, they are not necessarily the projection that leads to the greatest improvement in forecast accuracy. Therefore, referring to reconciliation via orthogonal projections as `optimal' is somewhat misleading since it does not have the optimality properties of some oblique projections, in particular MinT\@. {\color{blue} We now turn our attention to the way oblique projections can address both of these shortcomings} .

@@

One justification for using an orthogonal projection is that it leads to improved forecast accuracy in terms of the square root of the sum of squared errors of \emph{all} variables in the hierarchy. A clear shortcoming of this measure of forecast accuracy is that forecast errors in all series should not necessarily be treated equally. For example, in hierarchies, top-level series tend to have a much larger scale than bottom-level series. Even when two series are on a similar scale, series that are more predictable or less variable will tend to be down-weighted by simply aggregating squared errors. An even more sophisticated understanding may take the correlation between series into account.  {\color{blue} Finally, the context of the forecast problem itself may suggest a loss function, for instance in our tourism example an appropriate loss function may weight forecast errors by average tourism expenditure in each reason}. All of these considerations lead towards {\color{blue} a loss function $L_{\bm W}(\bm{y}_{t+h},\breve{\bm{y}}_{t+h|t})=||{\bm{y}_{t+h}-\breve{\bm{y}}_{t+h|t}}||_{\bm{W}^{-1}}$ where $||\bm{v}||_{\bm{W}^{-1}}=\bm{v}'\bm{W}^{-1}\bm{v}$, and $\bm{W}$ is a symmetric matrix assumed to be invertible.  The geometry defined by the norm $||.||_{\bm{W}^{-1}}$, will be referred to as the generalised Euclidean geometry with respect to $\bm{W}^{-1}$. }

One way to understand the generalied Euclidean geometry is that it is the same as Euclidean geometry when all vectors are first transformed by pre-multiplying by $\bm{W}^{-1/2}$. This leads to a transformed $\bm{S}$ matrix $\bm{S}^*=\bm{W}^{-1/2}\bm{S}$ and transformed $\hat{\bm{y}}$ and $\tilde{\bm{y}}$ vectors $\hat{\bm{y}}^*=\bm{W}^{-1/2}\hat{\bm{y}}$ and $\tilde{\bm{y}}^*=\bm{W}^{-1/2}\tilde{\bm{y}}$. {\color{blue} An projection of the form $\tilde{\bm{y}}=\bm{S}(\bm{S}'\bm{W}^{-1}\bm{S})^{-1}\bm{S}'\bm{W}^{-1}\hat{\bm{y}}$, is an} orthogonal projection in the transformed space since
\begin{align*}
  \tilde{\bm{y}}^*
    & = \bm{W}^{-1/2}\tilde{\bm{y}} \\
    & = \bm{W}^{-1/2}\bm{S}(\bm{S}'\bm{W}^{-1}\bm{S})^{-1}\bm{S}'\bm{W}^{-1}\hat{\bm{y}} \\
    & = \bm{S}^*(\bm{S}^{*'}\bm{S}^*)^{-1}\bm{S}^{*'}\hat{\bm{y}^*}.
\end{align*}

@@

\begin{theo}[General distance reducing property]\label{th:gdistred}
  If $\tilde{\bm{y}}_{t+h|t}=\bm{S}\bm{G}\hat{\bm{y}}_{t+h|t}$, where $\bm{G}$ is such that $\bm{S}\bm{G}$ is an orthogonal projection (in the generalised Euclidean sense) onto $\mathfrak{s}$ then:
  \begin{equation*}
    \|(\bm{y}_{t+h}-\tilde{\bm{y}}_{t+h|t})\|_{\bm{W}^{-1}}
      \le\
    |(\bm{y}_{t+h}-\hat{\bm{y}}_{t+h|t})\|_{\bm{W}^{-1}}.
  \end{equation*}
\end{theo}
\begin{proof}
  The proof is identical to the proof for Theorem~\ref{th:distred} but relies on the Generalised Pythagorean Theorem (applicable to Generalised Euclidean space) rather than the Pythagorean Theorem.
  \end{proof}

The implication of Theorem~\ref{th:gdistred} is that if the loss function is {\color{blue} given a priori as total sum of squared errors in some linearly transformed space,} then the projection matrix $\bm{S}(\bm{S}'\bm{W}^{-1}\bm{S})^{-1}\bm{S}'\bm{W}^{-1}$ is guaranteed to improve forecast accuracy over base forecasts.  {\color{blue}  This result does not depend on the covariance of forecast errors except in the special case where $\bm{W}$ is set to this covariance matrix.}

@@

\begin{theo}[MinT minimises expected loss]\label{th:gdistred}
    The usual MinT reconciliation method $\tilde{\bm{y}}=\bm{S}\left(\bm{S}'\bm{\Sigma}^{-1}\bm{S}\right)^{-1}\bm{S}'\bm{\Sigma}^{-1}\hat{\bm{y}}$ solves the optimisation problem in Equation~\ref{eq:GMinTObj} for any $\bm{W}$.
\end{theo}
\begin{proof}
	The loss function in Equation~\ref{eq:GMinTObj} is equivalent to Euclidean distance in the transformed space and can therefore be minimised by using the MinT method in this space.  The MinT method in the transformed space is give by 
	\begin{equation}
	\label{eq:MinTtran}
	\tilde{\bm{y}}^{*}=\bm{S}^{*}\left(\bm{S}^{*'}\bm{\Sigma}^{*-1}\bm{S}^{*}\right)^{-1}\bm{S}^{*'}\bm{\Sigma}^{*-1}\hat{\bm{y}}^{*}\,,
	\end{equation}
	where  $\bm{y}^{*}=\bm{W}^{-1/2}\bm{y}$, $\bm{S}^{*}=\bm{W}^{-1/2}\bm{S}$, $\hat{\bm{y}}^{*}=\bm{W}^{-1/2}\hat{\bm{y}}$ and
	\begin{align*}
	\bm{\Sigma}^*&=E\left[(\bm{y}^{*}-\hat{\bm{y}}^{*})(\bm{y}^{*}-\hat{\bm{y}}^{*})'\right]\\
	&=E\left[\bm{W}^{-1/2}(\bm{y}-\hat{\bm{y}})(\bm{y}-\hat{\bm{y}})'\bm{W}^{-1/2'}\right]\\
	&=\bm{W}^{-1/2}E\left[(\bm{y}-\hat{\bm{y}})(\bm{y}-\hat{\bm{y}})'\right]\bm{W}^{-1/2'}\\
	&=\bm{W}^{-1/2}\bm{\Sigma}\bm{W}^{-1/2'}
	\end{align*}
	Substituting the expressions for $\bm{\Sigma}^*$, $\bm{y}^{*}$ and $\hat{\bm{y}}^{*}$ into Equation~\ref{eq:MinTtran} yields, 
	\begin{eqnarray*}
	\bm{W}^{-1/2}\bm{y}=&\bm{W}^{-1/2}\bm{S}\left((\bm{W}^{-1/2}\bm{S})'(\bm{W}^{-1/2}\bm{\Sigma}\bm{W}^{-1/2'})^{*-1}\bm{W}^{-1/2}\bm{S}\right)^{-1}\\
	&(\bm{W}^{-1/2}\bm{S})'(\bm{W}^{-1/2}\bm{\Sigma}\bm{W}^{-1/2'})^{*-1}\bm{W}^{-1/2}\hat{\bm{y}}\\
	\end{eqnarray*}
\end{proof}

>>>>

In this section we discuss two sensible properties that can be achieved by reconciliation via orthogonal projection.
\begin{itemize}
\item The first is that reconciliation should adjust the base forecasts as little as possible; i.e. the base and reconciled forecasts should be `close'. 
\item The second is that reconciliation in some sense should improve forecast accuracy, or more loosely, that the reconciled forecast should be `closer' to the realised value targeted by the forecast.
\end{itemize}

To address the first of these properties we make the concept of closeness more concrete by considering the Euclidean distance between the base forecast $\hat{\bm{y}}_{t+h|t}$ and the reconciled forecast $\tilde{\bm{y}}_{t+h|t}$. A property of an orthogonal projection is that the distance between $\hat{\bm{y}}_{t+h|t}$ and $\tilde{\bm{y}}_{t+h|t}$ is minimal over any possible $\tilde{\bm{y}}\in\mathfrak{s}$. In this sense reconciliation via orthogonal projection {\color{blue} $\bm{G}=(\bm{S}'\bm{S})^{-1}\bm{S}'$} leads to the smallest possible adjustments of the base forecasts.  {\color{blue} If the Euclidean distance is interpreted as a loss function $L(\bm{u},\bm{v})=\|\bm{u}-\bm{v}\|$ and $\|.\|$, an orthogonal projection solves the optimisation problem $\underset{\bm{G}}{\min}\,L(\bm{S}\bm{G}\hat{\bm{y}}_{t+h|t},\hat{\bm{y}}_{t+h|t})$. This is a special case of the optimisation problem considered by \cite{NysEtAl2019}, the more general case will be covered in the next section.}

The {\color{blue} second property is concerned with the loss function $L({\bm{y}}_{t+h},\bar{\bm{y}}_{t+h|t})$ where $\bar{\bm{y}}_{t+h|t}$ is either a base or reconciled forecast.  More specifically, the aim is to reconcile so that $L({\bm{y}}_{t+h|t},\hat{\bm{y}}_{t+h})-L({\bm{y}}_{t+h},\tilde{\bm{y}}_{t+h|t})$ is guaranteed to be positive, that is} reconciled forecasts should always be closer to the target than base forecasts. {\color{blue} This is expressed as a minimax optimisation property by \cite{VanErven2015a} and was also} touched upon in Section 2.3 of \citet{WicEtAl2019}{\color{blue}, in both cases for a slightly different loss function}. Here we provide a new explicit proof of {\color{blue} the distance reducing property}. We do so first in the case {\color{blue} of the loss function defined here in terms of Euclidean distance} where the geometric intuition of the proof is clear and then generalise the result to {\color{blue} more general loss functions} Section~\ref{sec:oblique}

@@

In other ways, Theorem~\ref{th:distred} is weaker than perhaps often understood. First, when improvements in forecast accuracy are discussed in the context of the theorem, this refers to a very specific measure of forecast accuracy. In particular, this measure is the square root of the sum of squared forecast errors of \emph{all} variables in the hierarchy. Consequently, while forecast improvement is guaranteed for the hierarchy overall, {\color{blue} forecast errors are treated equally across all series}. Second, although orthogonal projections are guaranteed to improve on base forecasts, they are not necessarily the projection that leads to the greatest improvement in forecast accuracy. Therefore, referring to reconciliation via orthogonal projections as `optimal' is somewhat misleading since it does not have the optimality properties of some oblique projections. {\color{blue} We now turn our attention to the way oblique projections can address both of these shortcomings} .

@@

One justification for using an orthogonal projection is that it leads to improved forecast accuracy in terms of the square root of the sum of squared errors of \emph{all} variables in the hierarchy. A clear shortcoming of this measure of forecast accuracy is that forecast errors in all series should not necessarily be treated equally. For example, in hierarchies, top-level series tend to have a much larger scale than bottom-level series.  {\color{blue} The context of the forecast problem itself may suggest a loss function that weights series differently, for instance in our tourism example in Section~\ref{sec:EmpStudy} we will consider weighting forecast errors by average tourism expenditure in each region}.  An even more sophisticated understanding may take linear combinations of series into account. All of these considerations lead towards {\color{blue} a loss function $L_{\bm W}(\bm{y}_{t+h},\bar{\bm{y}}_{t+h|t})=||{\bm{y}_{t+h}-\breve{\bm{y}}_{t+h|t}}||_{\bm{W}}$ where $||\bm{v}||_{\bm{W}}=\bm{v}'\bm{W}\bm{v}$, and $\bm{W}$ is a symmetric matrix assumed to be invertible.  The geometry defined by the norm $||.||_{\bm{W}}$, will be referred to as the generalised Euclidean geometry with respect to $\bm{W}$. }

One way to understand the generalied Euclidean geometry is that it is the same as Euclidean geometry when all vectors are first transformed by pre-multiplying by $\bm{W}^{1/2}$, {\color{blue} where $\bm{W}=(\bm{W}^{1/2})'\bm{W}^{1/2}$}. This leads to a transformed $\bm{S}$ matrix $\bm{S}^*=\bm{W}^{1/2}\bm{S}$ and transformed $\hat{\bm{y}}$ and $\tilde{\bm{y}}$ vectors $\hat{\bm{y}}^*=\bm{W}^{1/2}\hat{\bm{y}}$ and $\tilde{\bm{y}}^*=\bm{W}^{1/2}\tilde{\bm{y}}$. {\color{blue} A projection of the form $\tilde{\bm{y}}=\bm{S}(\bm{S}'\bm{W}\bm{S})^{-1}\bm{S}'\bm{W}\hat{\bm{y}}$, is an} orthogonal projection in the transformed space since
\begin{align*}
  \tilde{\bm{y}}^*
    & = \bm{W}^{1/2}\tilde{\bm{y}} \\
    & = \bm{W}^{1/2}\bm{S}(\bm{S}'\bm{W}\bm{S})^{-1}\bm{S}'\bm{W}\hat{\bm{y}} \\
    &{\color{blue} = \bm{W}^{1/2}\bm{S}((\bm{W}^{1/2}\bm{S})'\bm{W}^{1/2}\bm{S})^{-1}(\bm{W}^{1/2}\bm{S})'\bm{W}^{1/2}\hat{\bm{y}}}\\
    & = \bm{S}^*(\bm{S}^{*'}\bm{S}^*)^{-1}\bm{S}^{*'}\hat{\bm{y}^*}.
\end{align*}

@@

\begin{theo}[General distance reducing property]\label{th:gdistred}
  If $\tilde{\bm{y}}_{t+h|t}=\bm{S}\bm{G}\hat{\bm{y}}_{t+h|t}$, where $\bm{G}$ is such that $\bm{S}\bm{G}$ is an orthogonal projection (in the generalised Euclidean sense) onto $\mathfrak{s}$ then:
  \begin{equation*}
    \|(\bm{y}_{t+h}-\tilde{\bm{y}}_{t+h|t})\|_{\bm{W}}
      \le\
    \|(\bm{y}_{t+h}-\hat{\bm{y}}_{t+h|t})\|_{\bm{W}}.
  \end{equation*}
\end{theo}
\begin{proof}
  The proof is identical to the proof for Theorem~\ref{th:distred} but relies on the Generalised Pythagorean Theorem (applicable to Generalised Euclidean space) rather than the Pythagorean Theorem.
  \end{proof}

The implication of Theorem~\ref{th:gdistred} is that if the loss function is of the form $L_{\bm{W}}$ {\color{blue} with some $\bm{W}$ given a priori} then the projection matrix $\bm{S}(\bm{S}'\bm{W}^{-1}\bm{S})^{-1}\bm{S}'\bm{W}^{-1}$ is guaranteed to improve forecast accuracy over base forecasts.  {\color{blue}  This result does not depend on the covariance of forecast errors except in general, although we note that MinT is a special case where $\bm{W}$ is set to this inverse of this covariance matrix.}

@@

\begin{theo}[MinT minimises expected loss]\label{th:gdistred}
    The usual MinT reconciliation method $\tilde{\bm{y}}=\bm{S}\left(\bm{S}'\bm{\Sigma}^{-1}\bm{S}\right)^{-1}\bm{S}'\bm{\Sigma}^{-1}\hat{\bm{y}}$ solves the optimisation problem in Equation~\ref{eq:GMinTObj} for any $\bm{W}$.
\end{theo}
\begin{proof}
	The loss function in Equation~\ref{eq:GMinTObj} is equivalent to Euclidean distance in the transformed space and can therefore be minimised by using the MinT method in this space.  The MinT method in the transformed space is give by 
	\begin{equation}
	\label{eq:MinTtran}
	\tilde{\bm{y}}^{*}=\bm{S}^{*}\left(\bm{S}^{*'}\bm{\Sigma}^{*-1}\bm{S}^{*}\right)^{-1}\bm{S}^{*'}\bm{\Sigma}^{*-1}\hat{\bm{y}}^{*}\,,
	\end{equation}
	where  $\bm{y}^{*}=\bm{W}^{-1/2}\bm{y}$, $\bm{S}^{*}=\bm{W}^{-1/2}\bm{S}$, $\hat{\bm{y}}^{*}=\bm{W}^{-1/2}\hat{\bm{y}}$ and
	\begin{align*}
	\bm{\Sigma}^*&=E\left[(\bm{y}^{*}-\hat{\bm{y}}^{*})(\bm{y}^{*}-\hat{\bm{y}}^{*})'\right]\\
	&=E\left[\bm{W}^{-1/2}(\bm{y}-\hat{\bm{y}})(\bm{y}-\hat{\bm{y}})'\bm{W}^{-1/2'}\right]\\
	&=\bm{W}^{-1/2}E\left[(\bm{y}-\hat{\bm{y}})(\bm{y}-\hat{\bm{y}})'\right]\bm{W}^{-1/2'}\\
	&=\bm{W}^{-1/2}\bm{\Sigma}\bm{W}^{-1/2'}
	\end{align*}
	Substituting the expressions for $\bm{\Sigma}^*$, $\bm{y}^{*}$ and $\hat{\bm{y}}^{*}$ into Equation~\ref{eq:MinTtran} yields, 
	\begin{align*}
	\bm{W}^{-1/2}\tilde{\bm{y}}=\bm{W}^{-1/2}\bm{S}&\left((\bm{W}^{-1/2}\bm{S})'(\bm{W}^{-1/2}\bm{\Sigma}\bm{W}^{-1/2'})^{*-1}\bm{W}^{-1/2}\bm{S}\right)^{-1}\\
	&(\bm{W}^{-1/2}\bm{S})'(\bm{W}^{-1/2}\bm{\Sigma}\bm{W}^{-1/2'})^{*-1}\bm{W}^{-1/2}\hat{\bm{y}}
	\end{align*}
	Rearranging and cancelling gives
	\begin{align*}
    \tilde{\bm{y}}&=\bm{S}\left((\bm{S})'(\bm{W}^{-1/2})'(\bm{W}^{1/2})'\bm{\Sigma}^{-1}\bm{W}^{1/2}\bm{W}^{-1/2}\bm{S}\right)^{-1}\bm{S}'(\bm{W}^{-1/2})'(\bm{W}^{1/2})'\bm{\Sigma}^{-1}\bm{W}^{1/2}\bm{W}^{-1/2}\hat{\bm{y}}\\
	&=\bm{S}\left((\bm{S})'\bm{\Sigma}^{-1}\bm{S}\right)^{-1}\bm{S}'\bm{\Sigma}^{-1}\hat{\bm{y}}\\
	\end{align*}
	which corresponds to the usual MinT method.
\end{proof}
